---
title: "Open Payments Data Cleaning"
author: "Peter Higgins"
date: "6/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(data.table)
library(dtplyr)
library(gt)
library(magrittr)
library(tigris)
library(zipcodeR)
library(scales)

specialty <- "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease"

# zip_table originally from https://gist.githubusercontent.com/erichurst/7882666/raw/5bdc46db47d9515269ab12ed6fb2850377fd869e/US%2520Zip%2520Codes%2520from%25202013%2520Government%2520Data
# 
zip_table <- read_csv("data_added/zip_table.csv")

# mapping - county name is in NAMELSAD column
county_sf <- tigris::counties()

### Read back in data from csv, convert to data.table

ds_full <- fread("dataset_full.csv")
```

## Test data.table for analysis

## Total pharma spend on pulmonary by year

1. Did this slow down during the Sunshine Act Era? Did Naming and Shaming Work?

2. What was the Effect of COVID?

First as a table, then as a plot.
```{r}
#  totals over 2013-2021, top 50
ds_full[ ,  .(total_usd = sum(Total_Amount_of_Payment_USDollars)),
    by = .(Program_Year)] %>% 
  tibble() ->  
pulm_annual_spend

# make pretty table
pulm_annual_spend %>% 
  gt() %>% 
  cols_label(Program_Year= "Year",
             total_usd = "Dollars") %>% 
  gt::tab_header(title = "Open Payments, Annual Pharma Spend on Pulmonary",
    subtitle = "2013-2021") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE)

# make plot
pulm_annual_spend %>% 
  ggplot(aes(x = Program_Year, y = total_usd)) +
  geom_point(color= 'red', size = 3) +
  geom_line() +
  scale_x_continuous(breaks = c(2013,2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021)) +
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  labs(x = 'Year' , y = 'US Dollars Paid',
       title = "Total Cumulative Paid to Pulmonologists by Pharma by Year",
       subtitle = "Millions of US Dollars",
       caption = "As Reported on CMS Open Payments")
```

While 2013 and 2014 may represent incomplete reporting, it appears that the overall trend was steady growth despite the Sunshine Act. This changed in 2020, when COVID had a much larger effect than the Sunshine Act, reducing pharma transfers by more than 50%, with a partial rebound in 2021.

### Data.table translation help
https://atrebas.github.io/post/2019-03-03-datatable-dplyr/#read-write-data 

## Total pharma spend on pulmonary by year by Category of General Spend (Nature of Payment field)

Perhaps some areas went down, like Food and Beverage, or Entertainment, in response to the Sunshine Act, even if overall spend went up.

one table, two graphs with different scales

```{r}
#  totals over 2013-2021, top 50
ds_full[ ,  .(total_usd = sum(Total_Amount_of_Payment_USDollars)),
    by = .(Program_Year, Nature_of_Payment_or_Transfer_of_Value)][order(Nature_of_Payment_or_Transfer_of_Value, Program_Year)] %>% 
  tibble() ->  
pulm_annual_cat_spend

# make pretty table
pulm_annual_cat_spend %>% 
  gt() %>% 
  cols_label(Program_Year= "Year",
             Nature_of_Payment_or_Transfer_of_Value = "Category",
             total_usd = "Dollars") %>% 
  gt::tab_header(title = "Open Payments, Annual Pharma Spend on Pulmonary",
    subtitle = "by Category, 2013-2021") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE)
```


### Now a plot of the big money

```{r}
# make plot - big spend first
pulm_annual_cat_spend %>% 
  filter(Nature_of_Payment_or_Transfer_of_Value %in% c("Compensation for services other than consulting, including serving as faculty or as a speaker at a venue other than a continuing education program", "Consulting Fee", "Food and Beverage", "Travel and Lodging", "Royalty or License")) %>% 
  ggplot(aes(x = Program_Year, y = total_usd, group = Nature_of_Payment_or_Transfer_of_Value)) +
  geom_point(aes(color= Nature_of_Payment_or_Transfer_of_Value), size = 3) +
  geom_line() +
  scale_x_continuous(breaks = c(2013,2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021)) +
  scale_y_continuous(labels = label_number(suffix = "M", scale = 1e-6)) +
  labs(x = 'Year' , y = 'US Dollars Paid',
       title = "Total Paid by Payment Category to Pulmonologists by Pharma by Year",
       subtitle = "Millions of US Dollars",
       caption = "As Reported on CMS Open Payments") +
  theme(legend.position = "bottom") +
  guides(col = guide_legend(ncol = 1))
```


### Now a plot of the small money

```{r}
# make plot - small spend 2nd
pulm_annual_cat_spend %>% 
  filter(Nature_of_Payment_or_Transfer_of_Value %in% c("Current or prospective ownership or investment interest", "Compensation for serving as faculty or as a speaker for a non-accredited and noncertified continuing education program", "Compensation for serving as faculty or as a speaker for a medical education program", "Charitable Contribution", "Education", "Entertainment", "Gift", "Grant")) %>% 
  ggplot(aes(x = Program_Year, y = total_usd, group = Nature_of_Payment_or_Transfer_of_Value)) +
  geom_point(aes(color= Nature_of_Payment_or_Transfer_of_Value), size = 3) +
  geom_line() +
  scale_x_continuous(breaks = c(2013,2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021)) +
  scale_y_continuous(labels = label_number(suffix = "K", scale = 1e-3)) +
  labs(x = 'Year' , y = 'US Dollars Paid',
       title = "Total Paid by Payment Category to Pulmonologists by Pharma by Year - Smaller Potatoes",
       subtitle = "Millions of US Dollars",
       caption = "As Reported on CMS Open Payments") +
  theme(legend.position = "bottom") +
  guides(col = guide_legend(ncol = 1))
```

Random non-CME speaking went up until COVID, consulting fees steadily up, Food & Bev up, with less COVID effect, while travel and royalties were severely affected by COVID. No clear effect of the Sunshine Act on Big-Ticket items.

For the smaller potatoe, Education went up steadily, even with COVID. Compensation for fake CME went down over time, grants slowly upward, and ownership was largely flat except for a 2018 spike. Gifts wmay have also gone down over time.


## Which Companies are spending the most, and on which products?

Top Drug Spend by Drug, Overall

```{r}
#  totals over 2013-2021, top 50
ds_full[ ,  .(total_usd = sum(Total_Amount_of_Payment_USDollars)),
    by = .(Name_of_Drug_or_Biological_or_Device_or_Medical_Supply_1)][order(-total_usd)] %>% 
  tibble() %>% 
  filter(Name_of_Drug_or_Biological_or_Device_or_Medical_Supply_1 != "") ->  
pharma_spend_per_drug

# make pretty table
pharma_spend_per_drug %>% 
  filter(total_usd > 1000000) %>% 
  gt() %>% 
  cols_label(Name_of_Drug_or_Biological_or_Device_or_Medical_Supply_1 = "Product",
        total_usd = "Dollars") %>% 
  gt::tab_header(title = "Open Payments, Total Pharma Spend on Pulmonary",
    subtitle = "by Product, 2013-2021") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE)
```

Top Table first, limited to a total of >=1M from 2013-2021

```{r}
#  totals over 2013-2021, top 50
ds_full[ ,  .(total_usd = sum(Total_Amount_of_Payment_USDollars)),
    by = .(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name, Name_of_Drug_or_Biological_or_Device_or_Medical_Supply_1)][order(-total_usd)] %>% 
  tibble() ->  
pharma_spend_by_drug

# make pretty table
pharma_spend_by_drug %>% 
  filter(total_usd > 1000000) %>% 
  gt() %>% 
  cols_label(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name = "PharmaCo",
     Name_of_Drug_or_Biological_or_Device_or_Medical_Supply_1 = "Product",
        total_usd = "Dollars") %>% 
  gt::tab_header(title = "Open Payments, Total Pharma Spend on Pulmonary",
    subtitle = "by Product, 2013-2021") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE)
```

## Total Spend by Company by Year

```{r}
#  totals over 2013-2021, top 50
ds_full[ ,  .(total_usd = sum(Total_Amount_of_Payment_USDollars)),
    by = .(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name, Program_Year)][order(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name, Program_Year)] %>% 
  tibble() %>% 
  filter(total_usd > 1000000) %>% arrange(-total_usd)->  
pharma_spend_by_year
```
## distinct companies over 1M in one year
Then a table of these

```{r}
pharma_spend_by_year %>% 
  distinct(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name) %>% 
  pull() ->
top_17_pharma

#pull top 17 companies that broke 1M at least once
ds_full[ Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name %in% top_17_pharma,  .(total_usd = sum(Total_Amount_of_Payment_USDollars)),
    by = .(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name, Program_Year)][order(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name, Program_Year)] ->
  top17_pharma_spend_annual

# tablulate these by year
top17_pharma_spend_annual %>% 
  gt() %>% 
  cols_label(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name = "PharmaCo",
     Program_Year = "Year",
        total_usd = "Dollars") %>% 
  gt::tab_header(title = "Open Payments, Big Pharma Spend on Pulmonary",
    subtitle = "by Company and Year, 2013-2021") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE)
```


### Now a plot of the spend by the big 17

```{r}
# now a plot
top17_pharma_spend_annual %>% 
  ggplot(aes(x = Program_Year, y = total_usd, group = Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name)) +
  geom_point(aes(color = Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name), size = 3) +
  geom_line() +
  scale_x_continuous(breaks = c(2013,2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021)) +
  scale_y_continuous(labels = label_number(suffix = "M", scale = 1e-6)) +
  labs(x = 'Year' , y = 'US Dollars Paid',
       title = "Total Paid by PharmaCos to Pulmonologists by Year",
       subtitle = "Millions of US Dollars",
       caption = "As Reported on CMS Open Payments") +
  theme(legend.position = "right") +
  guides(col = guide_legend(ncol = 1))
```

## Further Analysis

- how many payments to docs per year by company
- payments by doc within a company - histogram - KOL strategy?

## Now Top Individual Recipients

Not counting Royalties

```{r}
#  totals over 2013-2021, top 50
ds_full[Nature_of_Payment_or_Transfer_of_Value != "Royalty or License", ][ ,  .(total_usd = sum(Total_Amount_of_Payment_USDollars)),
    by = .(Physician_First_Name, Physician_Last_Name)][order(Physician_First_Name, Physician_Last_Name)] %>% 
  tibble() %>% 
  filter(total_usd > 1000000) %>% arrange(-total_usd)->  
docs_by_year
```

573 docs over 100K from 2013-2021
88 over 500K
23 over 1M

Table of Top 23

```{r}
docs_by_year %>% 
  gt() %>% 
  cols_label(Physician_First_Name =  "First Name",
     Physician_Last_Name = "Last Name",
        total_usd = "Dollars") %>% 
  gt::tab_header(title = "Open Payments, Big Pharma Spend on Pulmonary",
    subtitle = "by Doctor, 2013-2021") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE)
```

Of note, only one female

High number (over-representative) raised, educated outside the US on the top list?

What is the denominator - how many pulmonary docs in US, how many foreign-born, foreign-educated?

Academic vs non-academic?

Religious hospitals vs secular?

## Top by Royalty or License

Anyone over 10K
```{r}
#  totals over 2013-2021, top 50
ds_full[Nature_of_Payment_or_Transfer_of_Value == "Royalty or License", ][ ,  .(total_usd = sum(Total_Amount_of_Payment_USDollars)),
    by = .(Physician_First_Name, Physician_Last_Name)][order(Physician_First_Name, Physician_Last_Name)] %>% 
  tibble() %>% 
  filter(total_usd > 10000) %>% arrange(-total_usd)->  
docs_by_roy
```

Total of only 7 docs in 2013-2021.
 Lots for Lyntek Corp based on sepsis patents.
http://lyntekmedical.com 

Table
```{r}
docs_by_roy %>% 
  gt() %>% 
  cols_label(Physician_First_Name =  "First Name",
     Physician_Last_Name = "Last Name",
        total_usd = "Dollars") %>% 
  gt::tab_header(title = "Open Payments, Big Pharma Spend on Pulmonary",
    subtitle = "Royalties by Doctor, 2013-2021, Total > $10K") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE)
```

## Top Recipients from Top Companies
Which of the Top 17 Companies are spending the most on specific docs?


```{r}
#  totals over 2013-2021
library(dtplyr)
ds_full %>% 
  filter(Nature_of_Payment_or_Transfer_of_Value != "Royalty or License") %>% 
  filter(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name %in% top_17_pharma) %>% 
  group_by(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name, Physician_First_Name, Physician_Last_Name) %>% 
  summarise(total_usd = sum(Total_Amount_of_Payment_USDollars)) %>% 
  arrange(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name, -total_usd) %>% 
  filter(total_usd > 100000) ->  
docs_per_co

docs_per_co %>% 
    gt() %>% 
  cols_label(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name = "PharmaCo",
             Physician_First_Name =  "First Name",
     Physician_Last_Name = "Last Name",
        total_usd = "Dollars") %>% 
  gt::tab_header(title = "Open Payments, Big Pharma Spend on Pulmonary",
    subtitle = "Royalties by Doctor, 2013-2021, Total > $100K") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE)
```

## Top Recipients by Drug from Top Companies
Which of the Top 17 Companies are spending the most on specific docs for Specific Drugs?

Aka, Who is Team Symbicort vs Team Trelegy?

```{r}
#  totals over 2013-2021
library(dtplyr)
ds_full %>% 
  filter(Nature_of_Payment_or_Transfer_of_Value != "Royalty or License") %>% 
  filter(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name %in% top_17_pharma) %>% 
  group_by(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name, Name_of_Drug_or_Biological_or_Device_or_Medical_Supply_1, Physician_First_Name, Physician_Last_Name) %>% 
  summarise(total_usd = sum(Total_Amount_of_Payment_USDollars)) %>% 
  arrange(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name, Name_of_Drug_or_Biological_or_Device_or_Medical_Supply_1, -total_usd) %>% 
  filter(total_usd > 50000) ->  
docs_per_drug

docs_per_drug %>% 
    gt() %>% 
  cols_label(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name = "PharmaCo",
             Name_of_Drug_or_Biological_or_Device_or_Medical_Supply_1 = "Drug",
             Physician_First_Name =  "First Name",
     Physician_Last_Name = "Last Name",
        total_usd = "Dollars") %>% 
  gt::tab_header(title = "Open Payments, Big Pharma Spend on Pulmonary",
    subtitle = "Royalties by Doctor, 2013-2021, Total > $100K") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE)
```


## Who are the Hungry Hippos - most Food and Beverage $$

```{r}
#  totals over 2013-2021
library(dtplyr)
ds_full %>% 
  filter(Nature_of_Payment_or_Transfer_of_Value == "Food and Beverage") %>% 
  group_by(Physician_First_Name, Physician_Last_Name) %>% 
  summarise(total_usd = sum(Total_Amount_of_Payment_USDollars)) %>% 
  filter(total_usd > 20000) %>% 
    arrange(-total_usd) ->  
food_doc_dollars

food_doc_dollars %>% 
  mutate(Name = paste(Physician_First_Name, Physician_Last_Name)) %>% 
  ungroup() %>% 
  select(Name, total_usd) %>% 
    gt() %>% 
  cols_label(total_usd = "Dollars") %>% 
  gt::tab_header(title = "Who Are The Hungriest Hippos in Pulmonary",
    subtitle = "Food and Beverage by Doctor, 2013-2021, Total > $20K, N = 186") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE)
```


## Who Never Eats At Home - most Food and Beverage Payments

```{r}
#  totals over 2013-2021
library(dtplyr)
ds_full %>% 
  filter(Nature_of_Payment_or_Transfer_of_Value == "Food and Beverage") %>% 
  count(Physician_First_Name, Physician_Last_Name) %>% 
  filter(n > 900) %>% 
    arrange(-n) ->  
food_doc_meals

food_doc_meals %>% 
  mutate(Name = paste(Physician_First_Name, Physician_Last_Name)) %>% 
  ungroup() %>% 
  select(Name, n) %>% 
    gt() %>% 
  cols_label(n = "Meals") %>% 
  gt::tab_header(title = "Which Pulmonary Docs Never Eat At Home",
    subtitle = "Food and Beverage by Doctor, 2013-2021, N = 234\n More than 100 meals out per year - only 208 weeknights per year available to eat dinner out over 9 years - 9*208") 
```
