---
title: "open_payments"
author: "Peter Higgins"
date: "6/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(data.table)
library(knitr)
library(arrow)
library(duckdb)
library(gt)
```

## Read physician data

Read in the physician payment file and annual payment files from Open Payments.

```{r read}
provider_data <- fread(here("data_providers/OP_CVRD_RCPNT_PRFL_SPLMTL_P06302022.csv"))
```

There are 1.325 M providers listed.
16.9 % are non-physician Practitioners
"Covered Recipient Non-Physician Practitioner" vs
82.3% are "Covered Recipient Physician"

## Filter down and sum the data for Pulmonary and Critical Care

For 2013

```{r 2013}
data2013 <- fread(here("data_years/OP_DTL_GNRL_PGYR2013_P01222021.csv"))

data2013_pulm <- data2013[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2013_pulm_pmt <- data2013_pulm[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2013_pulm_pmt_short <- data2013_pulm_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2013) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2013_pulm)
rm(data2013_pulm_pmt)

data2013_cc <- data2013[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = c("Covered_Recipient_Profile_ID" = "Physician_Profile_ID")]

data2013_cc_pmt <- data2013_cc[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID"))

data2013_cc_pmt_short <- data2013_cc_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2013) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2013_cc)
rm(data2013_cc_pmt)
rm(data2013)

# show results
data2013_pulm_pmt_short
data2013_cc_pmt_short
```

For 2014

```{r 2014}
data2014 <- fread(here("data_years/OP_DTL_GNRL_PGYR2014_P01212022.csv"))

data2014_pulm <- data2014[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2014_pulm_pmt <- data2014_pulm[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2014_pulm_pmt_short <- data2014_pulm_pmt %>% 
   select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2013) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2014_pulm)
rm(data2014_pulm_pmt)

data2014_cc <- data2014[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2014_cc_pmt <- data2014_cc[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2014_cc_pmt_short <- data2014_cc_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2013) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2014_cc)
rm(data2014_cc_pmt)
#rm(data2014)

# show results
data2014_pulm_pmt_short
data2014_cc_pmt_short
```

For 2015

```{r 2015}
data2015 <- fread(here("data_years/OP_DTL_GNRL_PGYR2015_P01212022.csv"))

data2015_pulm <- data2015[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2015_pulm_pmt <- data2015_pulm[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2015_pulm_pmt_short <- data2015_pulm_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2015) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name)) 

rm(data2015_pulm)
rm(data2015_pulm_pmt)
rm(data2015_pulm)

data2015_cc <- data2015[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2015_cc_pmt <- data2015_cc[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2015_cc_pmt_short <- data2015_cc_pmt %>% 
   select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2015) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name)) 

rm(data2015_cc)
rm(data2015_cc_pmt)
rm(data2015)

# show results
data2015_pulm_pmt_short
data2015_cc_pmt_short
```

For 2016

```{r 2016}
data2016 <- fread(here("data_years/OP_DTL_GNRL_PGYR2016_P01212022.csv"))

data2016_pulm <- data2016[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2016_pulm_pmt <- data2016_pulm[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2016_pulm_pmt_short <- data2016_pulm_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2016) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name)) 

rm(data2016_pulm)
rm(data2016_pulm_pmt)

data2016_cc <- data2016[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2016_cc_pmt <- data2016_cc[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2016_cc_pmt_short <- data2016_cc_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2016) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name)) 

rm(data2016_cc)
rm(data2016_cc_pmt)
rm(data2016)

# show results
data2016_pulm_pmt_short
data2016_cc_pmt_short
```

For 2017

```{r 2017}
data2017 <- fread(here("data_years/OP_DTL_GNRL_PGYR2017_P01212022.csv"))

data2017_pulm <- data2017[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2017_pulm_pmt <- data2017_pulm[order(-total_pmt)] |>
   tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2017_pulm_pmt_short <- data2017_pulm_pmt %>% 
   select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2017) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2017_pulm)
rm(data2017_pulm_pmt)

data2017_cc <- data2017[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2017_cc_pmt <- data2017_cc[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 


data2017_cc_pmt_short <- data2017_cc_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2017) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2017_cc)
rm(data2017_cc_pmt)
rm(data2017)

# show results
data2017_pulm_pmt_short
data2017_cc_pmt_short
```

For 2018

```{r 2018}
data2018 <- fread(here("data_years/OP_DTL_GNRL_PGYR2018_P01212022.csv"))

data2018_pulm <- data2018[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2018_pulm_pmt <- data2018_pulm[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2018_pulm_pmt_short <- data2018_pulm_pmt %>% 
   select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2018) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2018_pulm_pmt)
rm(data2018_pulm)

data2018_cc <- data2018[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2018_cc_pmt <- data2018_cc[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2018_cc_pmt_short <- data2018_cc_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2018) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2018_cc_pmt)
rm(data2018_cc)
rm(data2018)

# show results
data2018_pulm_pmt_short
data2018_cc_pmt_short
```

For 2019

```{r 2019}
data2019 <- fread(here("data_years/OP_DTL_GNRL_PGYR2019_P01212022.csv"))

data2019_pulm <- data2019[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2019_pulm_pmt <- data2019_pulm[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2019_pulm_pmt_short <- data2019_pulm_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2019) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2019_pulm_pmt)
rm(data2019_pulm)

data2019_cc <- data2019[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2019_cc_pmt <- data2019_cc[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2019_cc_pmt_short <- data2019_cc_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2019) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2019_cc_pmt)
rm(data2019_cc)
rm(data2019)

# show results
data2019_pulm_pmt_short
data2019_cc_pmt_short
```

For 2020

```{r 2020}
data2020 <- fread(here("data_years/OP_DTL_GNRL_PGYR2020_P01212022.csv"))

data2020_pulm <- data2020[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2020_pulm_pmt <- data2020_pulm[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2020_pulm_pmt_short <- data2020_pulm_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2020) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2020_pulm_pmt)
rm(data2020_pulm)

data2020_cc <- data2020[Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Physician_Profile_ID"]

data2020_cc_pmt <- data2020_cc[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data, by = c("Physician_Profile_ID" = "Covered_Recipient_Profile_ID")) 

data2020_cc_pmt_short <- data2020_cc_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2020) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2020_cc_pmt)
rm(data2020_cc)
rm(data2020)

# show results
data2020_pulm_pmt_short
data2020_cc_pmt_short
```

For 2021

```{r 2021}
data2021 <- fread(here("data_years/OP_DTL_GNRL_PGYR2021_P06302022.csv"))

data2021_pulm <- data2021[Covered_Recipient_Type == "Covered Recipient Physician" & Covered_Recipient_Primary_Type_1 == "Medical Doctor" & Covered_Recipient_Specialty_1 == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Covered_Recipient_Profile_ID"]

data2021_pulm_pmt <- data2021_pulm[order(-total_pmt)] |>
   tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data) 

data2021_pulm_pmt_short <- data2021_pulm_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2021) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2021_pulm_pmt)
rm(data2021_pulm)

data2021_cc <- data2021[Covered_Recipient_Primary_Type_1 == "Medical Doctor" & Covered_Recipient_Specialty_1 ==  "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine", .(total_pmt = sum(Total_Amount_of_Payment_USDollars)), by = "Covered_Recipient_Profile_ID"]

data2021_cc_pmt <- data2021_cc[order(-total_pmt)] |>
  tibble() |>
  filter(total_pmt >50000) |>
  left_join(x = _, y = provider_data)

data2021_cc_pmt_short <- data2021_cc_pmt %>% 
  select(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name, Covered_Recipient_Profile_City, Covered_Recipient_Profile_State, total_pmt) %>% 
  mutate(year = 2021) |>
  filter(!is.na(Covered_Recipient_Profile_First_Name))

rm(data2021_cc_pmt)
rm(data2021_cc)
rm(data2021)

# show results
data2021_pulm_pmt_short
data2021_cc_pmt_short
```


now to clean up, bind rows, and print out a nice table.

```{r}

rm(provider_data)

# put all years together
compiled_pmt <- bind_rows(
          data2013_pulm_pmt_short,
          data2013_cc_pmt_short,
          data2014_pulm_pmt_short,
          data2014_cc_pmt_short,
          data2015_pulm_pmt_short,
          data2015_cc_pmt_short,
          data2016_pulm_pmt_short,
          data2016_cc_pmt_short,
          data2017_pulm_pmt_short,
          data2017_cc_pmt_short,
          data2018_pulm_pmt_short,
          data2018_cc_pmt_short,
          data2019_pulm_pmt_short,
          data2019_cc_pmt_short,
          data2020_pulm_pmt_short,
          data2020_cc_pmt_short,
          data2021_pulm_pmt_short,
          data2021_cc_pmt_short) %>% 
  arrange(-total_pmt)

#pulm only
compiled_pulm_pmt <- bind_rows(
          data2013_pulm_pmt_short,
          data2014_pulm_pmt_short,
          data2015_pulm_pmt_short,
          data2016_pulm_pmt_short,
          data2017_pulm_pmt_short,
          data2018_pulm_pmt_short,
          data2019_pulm_pmt_short,
          data2020_pulm_pmt_short,
          data2020_cc_pmt_short,
          data2021_pulm_pmt_short,
          data2021_cc_pmt_short) %>% 
  arrange(-total_pmt)

#how many unique docs?
compiled_pmt %>% 
  distinct(Covered_Recipient_Profile_First_Name, Covered_Recipient_Profile_Last_Name) %>% 
  count()

#cumulative revenue
compiled_pmt %>% 
  group_by(Covered_Recipient_Profile_First_Name,
           Covered_Recipient_Profile_Last_Name, 
           Covered_Recipient_Profile_City,
           Covered_Recipient_Profile_State) %>%   
  summarize(total = sum(total_pmt)) %>% 
  arrange(-total) -> 
cumul_rev
```

For 2013, experiment with {arrow} package

```{r}
# read in 4.17M rows
ds_2013 <- arrow::read_csv_arrow(here("data_years/OP_DTL_GNRL_PGYR2013_P01222021.csv"))

# filter for physician, pulm or CC
# down to 91K rows, 14 vars
ds_2013_pulm <- ds_2013 |> 
  filter(Covered_Recipient_Type == "Covered Recipient Physician" & Physician_Primary_Type == "Medical Doctor" & Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease") %>% 
  select(Physician_Profile_ID, Physician_First_Name, Physician_Middle_Name, Physician_Last_Name, Recipient_Primary_Business_Street_Address_Line1, Recipient_City, Recipient_State, Recipient_Zip_Code, Recipient_Country, Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name, Nature_of_Payment_or_Transfer_of_Value,  Name_of_Associated_Covered_Device_or_Medical_Supply1,
Total_Amount_of_Payment_USDollars, Date_of_Payment)

# WHO ate the most on the company dime in 2013?
ds_2013_pulm |>
  filter(Nature_of_Payment_or_Transfer_of_Value == "Food and Beverage") |>
  group_by(Physician_First_Name, Physician_Last_Name) %>% 
  summarize(total = sum(Total_Amount_of_Payment_USDollars)) |>
  arrange(-total) |>
  filter(total > 1000)

# WHAT companies paid the most to pulm/CC docs in 2013?
ds_2013_pulm |>
  group_by(Submitting_Applicable_Manufacturer_or_Applicable_GPO_Name) %>% 
  summarize(total = sum(Total_Amount_of_Payment_USDollars)) |>
  arrange(-total) 

# WHAT products paid the most to pulm/CC docs in 2013?
ds_2013_pulm |>
  group_by(Name_of_Associated_Covered_Device_or_Medical_Supply1) %>% 
  summarize(total = sum(Total_Amount_of_Payment_USDollars)) |>
  arrange(-total) 
```

For 2014 data, determine descriptive data by payment type
- min, mean, max, sd for each

```{r}
data2014 |>
  tibble() |>
   filter(Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease" | Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine") |>
  group_by(Nature_of_Payment_or_Transfer_of_Value) |>
  summarize(min = min(Total_Amount_of_Payment_USDollars),
            mean = mean(Total_Amount_of_Payment_USDollars),
            max = max(Total_Amount_of_Payment_USDollars),
            sd = sd(Total_Amount_of_Payment_USDollars)) 
    
```

For 2014 data, determine how many payments by payment type

```{r}
data2014 |>
  tibble() |>
   filter(Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease" | Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine") |>
  count(Nature_of_Payment_or_Transfer_of_Value) |>
  arrange(-n)
```


For 2014 data, what were the royalty or license payments (n =38)?
A total of 6 people (Mostly Ko-Pen Wang)
```{r}
data2014 |>
  tibble() |>
   filter(Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease" | Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine") |>
  filter(Nature_of_Payment_or_Transfer_of_Value == "Royalty or License") %>% 
  select(Physician_First_Name, Physician_Last_Name, Total_Amount_of_Payment_USDollars) |>
  print(n=Inf)

data2014 |>
  tibble() |>
   filter(Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease" | Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Critical Care Medicine") |>
  filter(Nature_of_Payment_or_Transfer_of_Value == "Royalty or License") %>% 
  group_by(Physician_First_Name, Physician_Last_Name) %>% 
  summarise(usd = sum(Total_Amount_of_Payment_USDollars)) %>% 
  arrange(-usd)
```

Try with arrow and all datasets

```{r}
ds_all <- arrow::open_dataset(here("data_years"), format = 'csv') %>% 
  arrow::to_duckdb()

# object.size()
# data2014 alone  5740326480 bytes ~ 5.7GB
# All years ~ 94M rows by 65 vars
# total for 2013-2021 is 47.7GB
# ds_all only 146792 bytes - mostly pointers to data

ds_all %>% 
  filter(Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease") |> 
group_by(Nature_of_Payment_or_Transfer_of_Value) |>
summarize(min = min(Total_Amount_of_Payment_USDollars),
median = median(Total_Amount_of_Payment_USDollars),
max = max(Total_Amount_of_Payment_USDollars),
sd = sd(Total_Amount_of_Payment_USDollars)) %>% 
  collect() %>% 
  gt::gt() %>% 
  gt::tab_header(title = "Open Payments Descriptive Data",
    subtitle = "Pulmonary 2016-2021") %>% 
  gt::fmt_number(columns = min:sd, decimals = 2, use_seps = TRUE) %>% 
  gt::gtsave("descriptive_pmt_types.html")

# timed with tictoc - 98 sec to process  
```

Try to get total payments by year to pulm/CC - is sunshine act working?
Total dollars spent.
Rising every year 2016-2019,
dropped >50% in 2020, then rose again in 2021

```{r}
tictoc::tic()
ds_all %>% 
  filter(Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease") |> 
  group_by(Program_Year) |>
  summarize(
  total_usd = sum(Total_Amount_of_Payment_USDollars)) %>% 
  collect() %>% 
  arrange(Program_Year) %>% 
  gt::gt() %>% 
  cols_label(Program_Year = "Year",
             total_usd = "Dollars") %>% 
  gt::tab_header(title = "Open Payments By Year",
    subtitle = "Pulmonary 2016-2021") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE) %>% 
  gt::gtsave("pmt_by_year.html")
tictoc::toc()
```

Top Ten Companies overall spend 2016-2021

```{r}
ds_all %>% 
  filter(Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease") |> 
  group_by(Applicable_Manufacturer_or_Applicable_GPO_Making_Payment_Name) |>
  summarize(
  total_usd = sum(Total_Amount_of_Payment_USDollars),
  total_pmt = count(Total_Amount_of_Payment_USDollars)) %>% 
  collect() %>% 
    slice_max(n = 25, order_by = total_usd) %>% 
  gt::gt() %>% 
  cols_label(Applicable_Manufacturer_or_Applicable_GPO_Making_Payment_Name = "Company",
             total_usd = "Dollars",
             total_pmt = "Number of Payments") %>% 
  gt::tab_header(title = "Open Payments, Top 25 Companies Overall",
    subtitle = "Pulmonary 2016-2021") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE) %>% 
  gt::fmt_number(columns = total_pmt, decimals = 0, use_seps = TRUE) %>% 
  gt::gtsave("company_usd_6year.html")
```

Top Ten companies by year

```{r}
tictoc::tic()
ds_all %>% 
  filter(Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease") |> 
  group_by(Program_Year, Applicable_Manufacturer_or_Applicable_GPO_Making_Payment_Name) |>
  summarize(
  total_usd = sum(Total_Amount_of_Payment_USDollars)) %>% 
  collect() %>% 
  ungroup() %>% 
  group_by(Program_Year) %>% 
    slice_max(n =10, order_by = total_usd) %>% 
  arrange(Program_Year, -total_usd) %>% 
  gt::gt() %>% 
  cols_label(Program_Year = "Year",
             Applicable_Manufacturer_or_Applicable_GPO_Making_Payment_Name = "Company",
             total_usd = "Dollars") %>% 
  gt::tab_header(title = "Open Payments, Top 10 Companies by Year",
    subtitle = "Pulmonary 2016-2021") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE) %>% 
  gt::gtsave("company_usd_by_year.html")
tictoc::toc()
```


GI top Ten Earners overall

```{r}
tictoc::tic()
ds_all %>% 
  filter(Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Gastroenterology") |> 
  select(Physician_First_Name, Physician_Last_Name, Total_Amount_of_Payment_USDollars) %>% 
  group_by(Physician_First_Name, Physician_Last_Name) |>
  summarize(
  total_usd = sum(Total_Amount_of_Payment_USDollars)) %>% 
  arrange(-total_usd) %>% 
  mutate(name = paste(Physician_First_Name, Physician_Last_Name)) %>% 
  ungroup() %>% 
  select(name, total_usd) %>% 
  slice_max(n = 12, order_by = total_usd) %>%
  collect() %>% 
  gt::gt() %>% 
  cols_label(name= "Name",
             total_usd = "Dollars") %>% 
  gt::tab_header(title = "Open Payments, Top 12 GI Payees Overall",
    subtitle = "GI 2016-2021, > $1.25 M Over 6 Years") %>% 
  gt::fmt_number(columns = total_usd, decimals = 2, use_seps = TRUE) %>% 
  gt::gtsave("gi_top.html")
tictoc::toc()
```


try with vroom - crashes - can load, but can't handle all the data in RAM

```{r}
files <- fs::dir_ls(here("data_years"))
files
df_all <- vroom::vroom(files)

df_all %>% 
  filter(Physician_Specialty == "Allopathic & Osteopathic Physicians|Internal Medicine|Pulmonary Disease") |> 
  group_by(Program_Year) |>
  summarize(
  total_usd = sum(Total_Amount_of_Payment_USDollars)) 
```

# clean up 2021 to 75 columns (format from 2106-2020)
```{r}
vroom::vroom("raw_data_2021_91vars/OP_DTL_GNRL_PGYR2021_P06302022.csv") %>% 
  # take out 16 extra columns
  select(-Covered_Recipient_NPI,
         -Covered_Recipient_Primary_Type_2,
         -Covered_Recipient_Primary_Type_3,
         -Covered_Recipient_Primary_Type_4,
         -Covered_Recipient_Primary_Type_5,
         -Covered_Recipient_Primary_Type_6,
         -Covered_Recipient_Specialty_2,
         -Covered_Recipient_Specialty_3,
         -Covered_Recipient_Specialty_4,
         -Covered_Recipient_Specialty_5,
         -Covered_Recipient_Specialty_6,
         -Associated_Device_or_Medical_Supply_PDI_1,
         -Associated_Device_or_Medical_Supply_PDI_2,
         -Associated_Device_or_Medical_Supply_PDI_3,
         -Associated_Device_or_Medical_Supply_PDI_4,
         -Associated_Device_or_Medical_Supply_PDI_5) %>% 
  #rename Covered_Recipient columns
  rename(Physician_Primary_Type = Covered_Recipient_Primary_Type_1,
        Physician_Profile_ID = Covered_Recipient_Profile_ID,
       Physician_First_Name =
         Covered_Recipient_First_Name,
       Physician_Middle_Name =
        Covered_Recipient_Middle_Name,
      Physician_Last_Name =
       Covered_Recipient_Last_Name,
      Physician_Name_Suffix =
      Covered_Recipient_Name_Suffix,
      Physician_Specialty = Covered_Recipient_Specialty_1,
      Physician_License_State_code1 = Covered_Recipient_License_State_code1, 
       Physician_License_State_code2 = Covered_Recipient_License_State_code2,
       Physician_License_State_code3 = Covered_Recipient_License_State_code3,
       Physician_License_State_code4 = Covered_Recipient_License_State_code4,
       Physician_License_State_code5 = Covered_Recipient_License_State_code5) %>% 
  write_csv("data_years/2021_75col.csv")
```

